---
title: "SEIR modelling"
output:
  html_document: default
  pdf_document: default
date: "2024-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("ggplot2")

source("solver.R")
source("model.R")
source("simulation.R")
source("plots.R")
```

## Modelling COVID-19 using a stochastic differential equation

### Equations

The SEIR model arises from the following assumptions:

$$
\begin{equation}
  S(t) + E(t) + I(t) + R(t) = N\quad  \forall t \\
  S \xrightarrow{\text{exposure rate }\beta} E \xrightarrow{\text{infection rate }\sigma} I \xrightarrow{\text{removal rate }\gamma}R
\end{equation}
$$

Consequently, we can determine the rate of transition and model this using a Poisson distribution.

$$
\begin{align}
  S \rightarrow E &\sim \text{Pois}\left( \frac{\beta}{N}S(t)I(t) \right) \\
  E \rightarrow I &\sim \text{Pois}\left( \sigma E(t)\right) \\
  I \rightarrow R &\sim \text{Pois}\left( \gamma I(t)\right)
\end{align}
$$

#### Estimating parameters for COVID-19

I approximated the values for $\beta$, $\sigma$ and $\gamma$ using the following:

-   Latent period of 2.5 days
-   Infectious period of 5.6 days
-   $R_0$ of 2.68

This resulted in the following parameters: $$
\begin{align}
\beta &= 0.47857 \\
\sigma &= 0.4 \\
\gamma &= 0.17857
\end{align}
$$

### SEIR solutions

#### Solving the equations

```{r covid_default_model}

parameters <- c(N = 10000, beta = 0.47857, sigma = 0.4, gamma = 0.17857)
state      <- c(S = 9990, E = 10, I = 0, R = 0)

out.default_solution <- simulate_seir(
  initial_value = state,
  params = parameters,
  start = 0,
  end = 100
)

plot_SEIR(out.default_solution, parameters, style="incidence")
```

### Generating multiple solutions (and a deterministic one) and comparing

```{r multi-plot}
out.deterministic <- simulate_seir(
  initial_value = state,
  params = parameters,
  start = 0,
  end = 100,
  stochastic = FALSE
)

out.stochastic <- replicate(100, simulate_seir(
  initial_value = state,
  params = parameters,
  start = 0,
  end = 100,
  stochastic = TRUE,
  simulate = FALSE
), simplify = FALSE)

plot_SEIR_incidences(out.deterministic, out.stochastic, parameters)
```

### Simulating using a real population

```{r}
parameters <- c(N = 100000, beta = 0.47857, sigma = 0.4, gamma = 0.17857)
state      <- c(S = 99999, E = 1, I = 0, R = 0)

out <- simulate_seir(
  initial_value = state,
  params = parameters,
  start = 0,
  end = 100
)

plot_SEIR(out, parameters, "combined")
plot_SEIR(out, parameters, "exploded")

# the histogram of events should exactly match the line plot
plot_SEIR(out, parameters, "incidence") + 
      geom_histogram(data=out$events, aes(x=time), binwidth=1, alpha=0.5)
```

Sanity check our results:

```{r}
results <- table(unlist(out$pop$.state))
results

unlist(tail(out$overview, n=1))
```

Calculate the generation time:

```{r}
# for each event, find the time when the original infector got infected
# original infection times
events <- out$events

events$infectee_exposure_time <- out$pop[events$infector, ]$E
events$gt <- events$time - events$infectee_exposure_time
```

```{r}
gt.back <- do.call(
  data.frame,
  aggregate(
    gt ~ time, data = events, FUN=function(x) {
      c(
        mean = mean(x),
        q1 = quantile(x, 0.25),
        q3 = quantile(x, 0.75),
        median = median(x),
        sd = sd(x)
      )
    }
  )
)
colnames(gt.back) <- c("time", "mean", "q1", "q3", "median", "sd")
gt.back
```

And forward generation:

```{r}
gt.fwd <- do.call(
  data.frame,
  aggregate(
    gt ~ infectee_exposure_time, data = events, FUN=function(x) {
      c(
        mean = mean(x),
        q1 = quantile(x, 0.25),
        q3 = quantile(x, 0.75),
        median = median(x),
        sd = sd(x)
      )
    }
  )
)
colnames(gt.fwd) <- c("time", "mean", "q1", "q3", "median", "sd")
gt.fwd
```

Graphing the results

```{r}
ggplot(gt.back, aes(x=time, y=mean)) + 
  geom_line() + 
  geom_errorbar(aes(ymin=mean - sd, ymax=mean + sd), width=.2,
                 position=position_dodge()) +
  labs(title="mean backwards generation time")
```

```{r}
ggplot(gt.fwd, aes(x=time, y=mean)) + 
  geom_line() + 
  geom_errorbar(aes(ymin=mean - sd, ymax=mean + sd), width=.2,
                 position=position_dodge()) +
  labs(title="mean backwards generation time")
```
